How much is too much? Item Response Theory procedures to shorten tests

### 2025-02-18 sis2025/LaTex+Package/styles/paper-sis-template.tex
update sis

\begin{equation}{eq:4pl}
	P(x_{pi}= 1| \theta_p, b_i, a_i, c_i, d_i) = c_i + (d_i -c_i) + \dfrac{\exp[a_i(\theta_p - b_i)]}{1 + \exp[a_i(\theta_p - b_i)]},
\end{equation}
where $\theta_p$ is the latent trait level of person $p$, $b_i$ is the location of the item on the latent trait (i.e., difficulty parameter, the higher the value, the higher the difficulty of the item), $a_i$ describes the ability of $i$ to discriminate between respondents with different latent trait levels (i.e.,discrimination parameter, the higher the value, the higher the discrimination ability of the item), and $c_i$ and $d_i$ describe the probability of observing a correct response when $\theta \to - \infty$ and $\theta \to +\infty$, respectively. 
When $\theta \to - \infty$, the probability of observing a correct response should tend to 0. Likewise, when $\theta \rightarrow +\infty$, the probability of observing a correct response should tend to 1. 
However, there might be instances where respondents with $\theta$ levels below the difficulty of the item, whom are hence expected not to respond correctly, might provide the correct response out of luck. The lucky guess parameter $c_i$ describes this probability, such that the probability of observing a correct response for $\theta \to - \infty$ tends to $c_i$ instead of 0. 
The same but inverse consideration applies when $\theta \to + \infty$. The $d_i$ parameter describes the probability of not endorsing the item given that the latent trait is above the location of the item, such that the probability of observing a correct response for $\theta \to + \infty$ tends to $d_i$ instead of 1.

@@

\begin{equation}{eq:iif}
	IIF_i = \dfrac{a^2[P(\theta)-c_i]^2[d_i - P(\theta)]^2}{(d_i-c_i)^2 P(\theta)Q(\theta)}.
\end{equation}
The informativeness of each item is strongly influenced by the location of the item on the latent trait with respect to a specific latent trait level $\theta$, its discriminativity, and the probability of lucky guess and careless error. In absence of lucky guess and careless error, the IIF reaches its maximum when the location of the item $b_i$ matches the latent trait level $\theta$, and it decreases as the distance between $b_i$ and $\theta$ increases. Moreover, the higher the discrimination of the item $a_i$, the more informative the item. However, when the lucky guess and careless error are taken into account, the informativeness of the item decreases, and, most importantly, the maximum of the IIF does not corresponds to the item location on the latent trait and it is generally lower.

@@

The two algorithms are based on the same principle, that is to reduce as much as possible the distance from a TIF target and the provisional TIF obtained from the items selected up to that point. 
At each iteration, the item is selected according to its ability of bridging the gap between the two TIFs. The algorithms stop when the last item considered for inclusion in the STF does not contribute to bring the provisional TIF closer to the TIF target, that is when the distance between the TIF target and the provisional TIF with the last considered item is equal to or greater than the distance between the TIF target and the provisional TIF without the last considered items (i.e., termination criterion).

@@

The parameters of the items in $B$ are defined through an $I \times 4$ matrix, where $I$ is the total number of items in $B$ ($||B||$) and the 4 columns contain the item parameters $b_i$, $a_i$, $c_i$, and $d_i$. 
Given that $B$ is the set of item without the effect of the tiredness (i.e., without careless parameters), $d_i = 1$, $\forall i \in \{1, \ldots, I\}$. To include the tiredness of the respondents, the vector of careless error parameters $d_i$ is modified with an exponential function, $d_i' = \exp(-\lambda r_i)$ (where $\lambda$ is the speed parameter that determines the steepness of the function and $r$ is the rank of the $i$-th item in the administration).
The set of items with $d_i'$ is denoted $B'$.

@@

At $k = 0$: $\forall i \in B$, $\text{IIF}_i$, $\text{TIF}^0(\theta) = 0 \, \forall \theta$, $Q^0 = \emptyset$. For $k \geq 0$,

\begin{enumerate}
	\item  $A^k = B' \setminus Q^k$ 
	\item $\forall i \in A^k$, $pTIF_{i}^k = \frac{\text{TIF}^k + \text{IIF}_{i}}{||Q^k||+1}$
	\item $i^* = \arg \min_{i \in A^k} |\text{TIF}_B - \text{pTIF}_i|$
	\item Termination criterion: $|\text{TIF}_B - \text{pTIF}_{i^*}| \geq |\text{TIF}_B - \text{TIF}^{k}|$: 
	\begin{itemize}
		\item FALSE:  $Q^{k+1} = Q^{k} \cup \{i^*\}$, $TIF^{k+1} = pTIF_{i^*}$, iterates 1-4 
		\item TRUE: Stop, %The item in $i^*$ does not contribute to reduce the distance from the TIF-target, hence: 
		$Q_{Frank} = Q^k$

@@

In this study, we operationalized the tiredeness of the individuals as the probability of committing careless error ($1- d_i$ QUESTA è LA PROBABILITà DI NON CARELESS ERROR DEVO SPIEGARE MEGLIO), such that the more the items in a test, the higher the probability of careless error. 
The careless error probability is instrinsically related to the rank of the item in the test, and it increases as the position of the item in the test increases. It increases exponentially as a function of the number of administered items, such that the first administered item has $d_i = 1$ (i.e., does not have a probability of observing an incorrect response given the latent trait level $\theta$).

@@

On average, Léon included more items in the STF than Frank did, $M_{Leon} = 6.75 \pm 3.06$, $M_{Frank} = 5.29 \pm 2.92$, $t = 3.45$, $df = 197.57$, $p < .001$. 
\color{red}
scrivere meglio 
Given that leon account for the tiredness of the respondents during the item selection, it also considers the measurement precision of the STF given that the respondents do get tired during the administration, while Frank does not. As such, while preventing the over inclusion of items in the STF, Léon is also able to account for the measurement prevision, in the attempt of balancing the number of items with the measurement precision.
\normalcolor

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/points-alogirthms}
	\caption{Distance from $TIF_{B}$ of $TIF_{B'}$ (green dots), $TIF_{Frank}$ (yellow triangles), and $TIF_{Leon}$ (brown squares) in each of the 100 replications. The horizontal green, yellow, and brown lines are the average distance from $TIF_B$ of $TIF_{B'}$, $TIF_{Frank}$, and $TIF_{Leon}$, respectively}
	\label{fig:points-alogirthms}
\end{figure}

>>>>

\begin{equation}\label{eq:4pl}
	P(x_{pi}= 1| \theta_p, b_i, a_i, c_i, d_i) = c_i + (d_i -c_i) + \dfrac{\exp[a_i(\theta_p - b_i)]}{1 + \exp[a_i(\theta_p - b_i)]},
\end{equation}
where $\theta_p$ is the latent trait level of person $p$, $b_i$ is the location of the item on the latent trait (i.e., difficulty parameter, the higher the value, the higher the difficulty of the item), $a_i$ describes the ability of $i$ to discriminate between respondents with different latent trait levels (i.e.,discrimination parameter, the higher the value, the higher the discrimination ability of the item), and $c_i$ and $d_i$ describe the probability of observing a correct response when $\theta \to - \infty$ and $\theta \to +\infty$, respectively. 
When $\theta \to - \infty$, the probability of observing a correct response should tend to 0. Likewise, when $\theta \rightarrow +\infty$, the probability of observing a correct response should tend to 1. 
However, there might be instances where respondents with $\theta$ levels below the difficulty of the item, whom are hence expected not to respond correctly, might provide the correct response out of luck. The lucky guess parameter $c_i$ describes this probability, such that the probability of observing a correct response for $\theta \to - \infty$ tends to $c_i$ instead of 0. 
The same but inverse consideration applies when $\theta \to + \infty$. The $d_i$ parameter describes the probability of not endorsing the item given that the latent trait is above the location of the item, such that the probability of observing a correct response for $\theta \to + \infty$ tends to $d_i$ instead of 1.

@@

\begin{equation}\label{eq:iif}
	IIF_i = \dfrac{a^2[P(\theta)-c_i]^2[d_i - P(\theta)]^2}{(d_i-c_i)^2 P(\theta)Q(\theta)}.
\end{equation}
The informativeness of each item is strongly influenced by the location of the item on the latent trait with respect to a specific latent trait level $\theta$, its discriminativity, and the probability of lucky guess and careless error. In absence of lucky guess and careless error, the IIF reaches its maximum when the location of the item $b_i$ matches the latent trait level $\theta$, and it decreases as the distance between $b_i$ and $\theta$ increases. Moreover, the higher the discrimination of the item $a_i$, the more informative the item. However, when the lucky guess and careless error are taken into account, the informativeness of the item decreases, and, most importantly, the maximum of the IIF does not corresponds to the item location on the latent trait and it is generally lower.

@@

In this study, we operationalized the tiredeness of the individuals as the probability of committing careless error, such that the more the items in a test, the higher the probability of careless error on the $i$-th item. 
The careless error probability is instrinsically related to the rank of the item in the test, and it increases exponentially as a function of the number of administered items.

The two algorithms are based on the same principle of reducing as much as possible the distance between a TIF-target and a provisional TIF (pTIF) obtained from the items selected up to that point. 
At each iteration, the item is included in the STF according to its ability of bridging the gap between the two TIFs. 
The algorithms stop when the last item considered for inclusion in the STF does not contribute to bring the pTIF closer to the TIF-target, that is when the distance between the TIF-target and the pTIF with the last considered item is equal to or greater than the distance between the TIF-target and the pTIF without the last considered items (i.e., termination criterion).

@@

The parameters of the items in $B$ are defined through an $I \times 4$ matrix, where $I$ is the total number of items in $B$ ($||B||$, where $||X||$ denotes the cardinality of set $X$) and the 4 columns contain the item parameters $b_i$, $a_i$, $c_i$, and $d_i$. 
Given that $B$ is the set of item without the effect of the tiredness (i.e., without careless parameters), $d_i = 1$, $\forall i \in \{1, \ldots, I\}$. To include the tiredness of the respondents, the vector of careless error parameters $d_i$ is modified with an exponential function, $d_i' = \exp(-\lambda r_i)$ (where $\lambda$ is the speed parameter that determines the steepness of the function and $r$ is the rank of the $i$-th item in the administration).
The set of items with $d_i'$ is denoted $B'$.

Since the TIF increases as the number of items in a test increases, the comparison between $TIF_B$ and the TIF of the STF obtained by the item selections provided by the two algorithms is based on the mean TIF (i.e., the TIF divided by the number of items in the STF). Nonetheless, in what follows the mean TIF will be simply referred to as $TIF_x$, with $x \in \{B, B', \text{Frank}, \text{Leon}\}$.

@@

\color{red}
Non è propriamente così, frank calcola tutte le IIF all'inziio considerando i parametri degli item in $B'$ e se le porta avanti ma quindi dovrei creare un altro oggetto che contiene tutte le iif i? ci provo tanto ormai è tutto una merda 
\normalcolor

At $k = 0$: $\mathbf{IIF} = $ $\forall i \in B$, $\text{IIF}_i$, $\text{TIF}^0(\theta) = 0 \, \forall \theta$, $Q^0 = \emptyset$. For $k \geq 0$,

\begin{enumerate}
	\item  $A^k = B \setminus Q^k$ 
		\color{red}
	\item $\forall i \in A^k$, $pTIF_{i}^k = \frac{\text{TIF}^k + \text{IIF}_{i}}{||Q^k||+1}$
	\normalcolor
	\item $i^* = \arg \min_{i \in A^k} |\text{TIF}_B - \text{pTIF}_i|$
	\item Termination criterion: $|\text{TIF}_B - \text{pTIF}_{i^*}| \geq |\text{TIF}_B - \text{TIF}^{k}|$: 
	\begin{itemize}
		\item FALSE:  $Q^{k+1} = Q^{k} \cup \{i^*\}$, $TIF^{k+1} = pTIF_{i^*}$, iterates 1-4 
		\item TRUE: Stop, %The item in $i^*$ does not contribute to reduce the distance from the TIF-target, hence: 
		$Q_{Frank} = Q^k$

@@

@@

On average, Léon included more items in the STF than Frank did, $M_{Leon} = 6.75 \pm 3.06$, $M_{Frank} = 5.29 \pm 2.92$, $t = 3.45$, $df = 197.57$, $p < .001$. 
\color{red}
Given that leon account for the tiredness of the respondents during the item selection, it also considers the measurement precision of the STF given that the respondents do get tired during the administration, while Frank does not. As such, while preventing the over inclusion of items in the STF, Léon is also able to account for the measurement precision, in the attempt of balancing the number of items with the measurement precision.
\normalcolor

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\linewidth]{img/box-plot-alogirthms}
	\caption{Distance from $TIF_{B}$ of $TIF_{B'}$ (green dots), $TIF_{Frank}$ (yellow triangles), and $TIF_{Leon}$ (brown squares) in each of the 100 replications. The horizontal green, yellow, and brown lines are the average distance from $TIF_B$ of $TIF_{B'}$, $TIF_{Frank}$, and $TIF_{Leon}$, respectively}
	\label{fig:points-alogirthms}
\end{figure}
